---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into multiple options: under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus. How many loans are in each group?

**A1** btw_100k_500k there are 4152, btw_10k_50k there are 15372, btw_500k_1m there are 540,btw_50k_100k there are 3978, more_than_1m there are 356, under_10k there are 9310

```{r}
wv_ppp_loan_amount_categories2 <- west_virginia_ppp %>%
  select(amount) %>%
  mutate(label = case_when(
    amount < 10000 ~ "under_10k", 
    amount >= 10000 & amount < 50000 ~ "btw_10k_50k",
    amount >= 50000 & amount < 100000 ~ "btw_50k_100k",
    amount >= 100000 & amount < 500000 ~ "btw_100k_500k",
    amount >= 500000 & amount < 1000000 ~ "btw_500k_1m",
    amount >= 1000000 ~ "more_than_1m")) %>% 
  group_by(label) %>% 
  summarise(total = n())

```

**Q2** Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Harvest Small Business Finance, LLC

```{r}
west_virginia_ppp %>% 
  filter(undisbursed_amount > 0 ) %>%  
  group_by(servicing_lender_name) %>% 
  summarize(total = n()) %>% 
  arrange(desc(total))
```

**Q3** How would you describe the top 10 list of those lenders? What do they have in common or not? (You may want to look up some of the top ones.)
**A3** Most of the top ten lenders work in banking and finance. 

**Q4** Which month in 2021 had the most loans approved?
**A4** April

```{r}
west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>% 
  group_by(Month) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))
```
